---
title: "Figure for cohort overlap and simulations."
output: html_notebook
---
Here, creating the simulation based figure
Note that previous sims were incorrect in that they didn't quite use the sd specifications in the right way. We should be able to get this to work
## Update the paths below to reflect the new locations, things have been mooved about!"
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/simulation_processing_tools.R")
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/plot_functions.R")
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/factorization_methods.R")
source("/data/abattle4/aomdahl1/gene_utils/qqplot.R")
library(data.table)
library(magrittr)
library(dplyr)
library(stringr)
library(ggplot2)
library(RColorBrewer)
library(NatParksPalettes)
library(ggbeeswarm)
library(ggpubr)

```

first, the null simulations based on GWAS
Updated Oct 2024
```{r}
#this one was made in the old way, so just loaded the files directly in
load("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/null_matrix_spearman/N30000_multi-overlap_SD-1_h2-0-0_bmi-like.SCALED_Aug2024/null_covarfactorization.summaries.RData")
OLD_VERSION=FALSE
if(OLD_VERSION)
{
  #The format has changed a little- I had to do the whitened svd separately
load("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/null_matrix_spearman/N30000_multi-overlap_SD-1_h2-0-0_bmi-like.SCALED_Aug2024/null_covarfactorization.summaries.WHITENED_PCA.RData")
save.list$elbow.point <- elbow.point
save.list$flashr.runs <- flashr.runs
save.list$gd.point <- gd.point
save.list$GLEANR.runs.covar <- GLEANR.runs.covar
save.list$GLEANR.runs.std <- GLEANR.runs.std
save.list$sim.pca.alt <- sim.pca.alt
save.list$sim.pca.std <- sim.pca.std
save.list$kaiser.point <- kaiser.point
save.list.joined <- save.list
#rerun on flash:
}
save.list.joined <- save.list
load("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/null_matrix_spearman/N30000_multi-overlap_SD-1_h2-0-0_bmi-like.SCALED_Aug2024/null_covarfactorization.summaries.flash_SE_update.RData")
save.list.joined$flashr.runs.se <- save.list$flashr.runs.se
#temporary hack while bringing things up to the newest organization fo the software
tables.alt <- tableSimResultsFromList(save.list.joined)
#}else
#{
#  #save.list$flashr.runs <- save.list$flashr.runs.se
#  tables.alt <- tableSimResultsFromList(save.list)
#}

factors.by.method.long <- tables.alt$long
```
Now we can work as in the previous ones:
```{r}
#Do by percent of overlapping samples
factors.by.method.long %<>% 
  mutate("overlap_percent" = overlap_score/90000*100) %>%  
  mutate("overlap_bins" = cut_interval(overlap_percent, n=4)) %>% 
  filter(method != "whitened_gd.point", method != "gd.point")

```
Build specific color pallette
Choices here- dropping SVD-beta
```{r}
map.to.new <- data.frame(
  "method"=unique(factors.by.method.long$method),
"method_label"=c("GLEANR", "GLEANR-U", "SVD-adj (elbow)", 
               "SVD-adj (avg)","flash", "SVD (elbow)", "SVD (avg)"))

#ordered
"new.method"=c("SVD (avg)", "SVD (elbow)", "SVD-adj (avg)","SVD-adj (elbow)","flash", "GLEANR-U","GLEANR")

full.list=c("SVD", "SVD (avg)", "SVD (elbow)","SVD-adj", "SVD-adj (avg)","SVD-adj (elbow)",
            "flash","flash (Z)","flash (Kronecker)","FactorGo","GLEANR-U","GLEANR")
#Trying some color things....
#flash.cols <- natparks.pals(name="Acadia",n=3,type="discrete")
#pca.unadj.cols <- natparks.pals(name="Glacier",n=3,type="discrete")
#pca.adj.cols <- natparks.pals(name="CapitolReef",n=3,type="discrete")
#GLEANR.cols <- natparks.pals(name="BryceCanyon",n=2,type="discrete")
#factorgo.cols <- natparks.pals(name="Arches",n=1,type="discrete")
#

flash.cols <- brewer.pal(5,"Greens")[3:5]
pca.unadj.cols <- brewer.pal(3,"Blues")
pca.adj.cols <- brewer.pal(8,"BuPu")[c(5,6,8)]
pca.adj.cols <- brewer.pal(3,"Purples")
GLEANR.cols <- brewer.pal(2,"Oranges")[2:3]
factorgo.cols <- brewer.pal(5,"PuRd")[4]



all.colors <- c(pca.unadj.cols,pca.adj.cols, flash.cols, factorgo.cols,GLEANR.cols )
color.map.pub <- data.frame("method_label"=full.list, "color" = all.colors)
```

Map
```{r}
factors.by.method.long.sub.plot <- left_join(factors.by.method.long, map.to.new, by="method") %>% left_join(., color.map.pub, by="method_label")%>%  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% arrange(method_label)

sub.pub <- color.map.pub %>% mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% filter(method_label %in% factors.by.method.long.sub.plot$method_label)

```
Plot
```{r}

myColors <- color.map.pub$color
names(myColors) <- levels(factors.by.method.long.sub.plot$method_label)
#names(myColors) <- unique(factors.by.method.long.sub.plot$method_label)
colScale <- scale_colour_manual(name = "Method",values = myColors)
fillScale <- scale_fill_manual(name = "Method",values = myColors)


hum_names <- as_labeller(
     c(`[0,25]` = "0-25%", `(25,50]` = "25-50%",`(50,75]` = "50-75%", 
       `(75,100]` = "75-100%"))

gwas.based <- ggplot(factors.by.method.long.sub.plot, aes(x=method_label,y=as.numeric(value), fill=method_label))+
  geom_violin(aes(fill = method_label)) +theme_bw(15) +  geom_jitter(alpha = 0.1,aes(color=method_label))+
 fillScale + facet_wrap(~overlap_bins, nrow = 1,labeller = hum_names, strip.position = "top") +  theme(strip.background =element_rect(fill="white")) + colScale + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),axis.title.x = element_blank()) + ylab(expression("# estimated factors" ~ hat(K))) 
 # ggtitle("Low levels of cohort overlap sufficient for common methods to detect spurious components") + theme(plot.title = 
#element_text(size=10,color="dimgrey"))  + xlab("% overlapping samples")  +
#+ ylab(expression( hat(K) ))
```
Great, Im' actually pretty happy with that.
##################################
Next, the factor based simulations. Moment of truth!!!

Unscaled flavor
```{r, eval = FALSE}
no.covar.unscaled <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/no_overlap/", scale=FALSE)  %>% mutate("Covar"=  "none")

oneb.covar.unscaled  <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/1b_overlap/", scale=FALSE) %>%   mutate("Covar"=  "1b")
twob.covar.unscaled  <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/2b_overlap/", scale=FALSE)  %>%   mutate("Covar"=  "2b")

all.sims.joined.unscaled <- rbind(no.covar.unscaled,oneb.covar.unscaled,twob.covar.unscaled)
```

Scaled flavor (preferred, more comparable across since most methods autoscale)
Go with this- unscaled isn't really fair.
```{r}
no.covar <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/no_overlap/")  %>% 
  mutate( "Covar"=  "none")

oneb.covar <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/1b_overlap/")  %>% 
  mutate("Covar"=  "1b")
twob.covar <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/2b_overlap/")  %>% 
  mutate("Covar"=  "2b")

all.sims.joined.scaled <- rbind(no.covar,oneb.covar,twob.covar)

```



```{r}
all.sims.joined.scaled$N <- factor(all.sims.joined.scaled$N, levels = c("5000","10000","50000","1e+05","2e+05","mixed"))
#all.sims.joined.unscaled$N  <- factor(all.sims.joined.unscaled$N, levels = c("5000","10000","50000","1e+05","2e+05","mixed"))
#Methods:

full.list=c("SVD", "SVD (avg)", "SVD (elbow)","SVD-adj", "SVD-adj (avg)","SVD-adj (elbow)",
            "flash","flash (Z)","flash (Kronecker)","FactorGo","GLEANR-U","GLEANR")

drop.methods <- c()

#add fakes to make sure all methods are included.....
fake.row.svd.elbow <- all.sims.joined.scaled[1,]; fake.row.svd.elbow$method="SVD (elbow)";  fake.row.svd.elbow[,8:35] <- NA
fake.row.svd.adjelbow <- all.sims.joined.scaled[1,]; fake.row.svd.adjelbow$method="SVD-adj (elbow)";  fake.row.svd.adjelbow[,8:35] <- NA
fake.row.svd.adj_avg <- all.sims.joined.scaled[1,]; fake.row.svd.adj_avg$method="SVD-adj (avg)";  fake.row.svd.adj_avg[,8:35] <- NA
all.sims.joined.scaled <- rbind(all.sims.joined.scaled, fake.row.svd.elbow,fake.row.svd.adjelbow,fake.row.svd.adj_avg)
#all.sims.joined.unscaled <- rbind(all.sims.joined.unscaled, fake.row)

name.map = data.frame("method_label"=c("SVD", "flash (Z)", "flash", "flash (Kronecker)","SVD (avg)", "GLEANR", "GLEANR-U", "SVD-adj", "SVD (beta)", "FactorGo", "SVD (elbow)","SVD-adj (elbow)","SVD-adj (avg)"), "method" = unique(all.sims.joined.scaled$method))

level_order=name.map$method_label
all.sims.joined.scaled %<>% left_join(.,name.map, by="method") %>% 
  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% arrange(method_label) %>% filter(method != "SVD_beta",method_label != "flash (Z)")

#all.sims.joined.unscaled %<>% left_join(.,name.map, by="method") %>% 
#  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% arrange(method_label) %>% filter(method != #"SVD_beta",method_label != "flash (Z)")

```
Numerical checks- where do we outperform other methods?
```{r}
#Get for U- my claim holds true!
all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianU) %>% filter(Covar == "1b") %>% group_by(N) %>% slice_max(medianU)

all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianU) %>% filter(Covar == "2b") %>% group_by(N) %>% slice_max(medianU)

all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianU) %>% filter(Covar == "none") %>% group_by(N) %>% slice_max(medianU)
```

Claim for V?
```{r}
#Get for V- my claim holds true!
all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianV) %>% filter(Covar == "1b") %>% group_by(N) %>% slice_max(medianV)

all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianV) %>% filter(Covar == "2b") %>% group_by(N) %>% slice_max(medianV)

all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianV) %>% filter(Covar == "none") %>% group_by(N) %>% slice_max(medianV)
```

Make the plot
Decision to drop blash Z, too compli
```{r}
##Scaled
pop_names <- as_labeller(
     c(`5000` = "N=5k", `10000` = "N=10k",`50000` = "N=50k", 
       `1e+05` = "N=100k",`2e+05` = "N=200k", "mixed"="N=[10K,200K]",
     `1b`="1-block", `2b`="2-block", `none`="No overlap"))

# Supplemental figure
drop.methods <- c("flash (Z)", "SVD (elbow)","SVD (elbow)","SVD-adj (elbow)","SVD-adj (avg)", "SVD (avg)")

v.sim <- ggplot(all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)), aes(x = method_label, y = yuan_V, fill = method_label)) + geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab(bquote(R^2 ~ "with simulated V")) + fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + ggtitle("Performance across simulations with respect to V")
v.leg <- ggpubr::get_legend(v.sim)

#Chat GPT helped me with this one.
u.sim.for.legend <- ggplot(all.sims.joined.scaled %>% filter(method_label != "flash (Z)"), 
                aes(x = method_label, y = yuan_U, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw() +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + theme(legend.text=element_text(size=12))
#bottom.legend = get_plot_component(plot, 'guide-box-left', return_all = TRUE)
bottom.legend <- get_legend(u.sim.for.legend)
drop.methods <- c("flash (Z)", "SVD (elbow)","SVD (elbow)","SVD-adj (elbow)","SVD-adj (avg)", "SVD (avg)")
u.sim <- ggplot(all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = yuan_U, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + theme(legend.text=element_text(size=12)) + theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt")) + ggtitle("Covariance adjustment improves reconstruction of simulated genetic factors") + theme(plot.title = element_text(size=10,color="dimgrey"))  
```

some arrows to go along with it (kkeep tweaking)
```{r}
library(ggplot2)
blurb = "Better recovery of U"
improvement <- ggplot() + xlim(c(-0.1,0.1)) + ylim(c(-1,0.5))+
  annotate("text", x = -0.05, y = -0.1, size=3.5, label = blurb, angle=90) + 
  theme_nothing() + annotate("segment", 
           x = 0, xend = 0, y = -0.9, yend = 0,
           arrow=arrow())


decline <- ggplot() + xlim(c(-0.1,0.1)) + ylim(c(-1,1.5))+
  annotate("text", x = -0.05, y = -0.1, size=3.5, label = "Worse recovery of U", angle=90) + 
  theme_nothing() + annotate("segment", 
           x = 0, xend = 0, y = 1, yend = -0.2,
           arrow=arrow())

arrows <- cowplot::plot_grid(plotlist =list(improvement,ggplot()+theme_nothing(), decline), ncol=1,rel_heights = c(1,1.3,1))
blank.arrows <- ggplot()+theme_nothing()
```

Results basically look the same, kinda. Scaled is more favorable for us, and I think its more fair all around.
#### Supplemenetal figure- looking at AUC, Kappa, F1:
Not doing Kappa or F1 here- Kappa just looks at agreement, when no known truth. F1 we don't have the precisoin/recall numbers close at hand.
```{r}

kappa.v  <- ggplot(all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = joint_cb_kappa_v, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(kappa ~ "for simulated V")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + theme(legend.text=element_text(size=12)) + theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt"))  + ggtitle("Classification (0/+/-) across simulations with respect to V")

kappa.u <- ggplot(all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = joint_cb_kappa_u, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(kappa ~ "for simulated U")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + theme(legend.text=element_text(size=12)) + theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt"))  + ggtitle("Classification (0/+/-) across simulations with respect to U")

cowplot::plot_grid(v.sim + theme(legend.position = "none"),kappa.v, kappa.u,v.leg,ncol=1,rel_heights = c(1,1,1,0.3)) 
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/suppl6_alternate_sim_metrics.svg",
      width = 8,height=12,units="in")
```



### Heatmaps of block structure
All right. The only figures remaining here are the- adjcacent 
```{r}
oneb.version <- fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/1b_overlap/V101_U101_MAF-mix_eur_N-10000_RHO-1b_high_mixed_p_No-1b_high_no/sim1.c_matrix.txt")

oneb.plot <- plotCorrelationHeatmap(as.matrix(oneb.version),typin = "none") + theme_minimal()+
  theme(axis.text.x=element_blank(),axis.text.y = element_blank(), legend.position = "none",
        axis.ticks = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=3)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(plot.margin = unit(c(-0.5,-1,-0.5,-1), "cm")) + coord_fixed()
  
```

```{r}
twob.version <- fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/2b_overlap/V101_U101_MAF-mix_eur_N-10000_RHO-2b_mid_mixed_p_No-2b_high_no/sim1.c_matrix.txt")

twob.plot <- plotCorrelationHeatmap(as.matrix(twob.version),typin = "none") + theme_minimal()+
  theme(axis.text.x=element_blank(),axis.text.y = element_blank(), legend.position = "none",
        axis.ticks = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=3)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(plot.margin = unit(c(-0.5,-1,-0.5,-1), "cm"))+ coord_fixed()
```

```{r}
nob.version <- fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/no_overlap/V101_U101_MAF-mix_eur_N-10000_RHO-none_No-none/sim1.c_matrix.txt")

nob.plot <- plotCorrelationHeatmap(as.matrix(nob.version),typin = "none") + theme_minimal()+
  theme(axis.text.x=element_blank(),axis.text.y = element_blank(), legend.position = "none",
        axis.ticks = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=3)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(plot.margin = unit(c(-0.5,-1,-0.5,-1), "cm")) + coord_fixed()
```

```{r}
blockheader <- ggdraw() +
  draw_label("C in simulation",size = 9,color = "dimgrey") + theme(plot.margin = unit(c(0,0,-1,0), "cm"))

cowplot::plot_grid(blockheader, 
                                 oneb.plot, twob.plot, nob.plot, ggplot() + geom_blank() + theme_classic(),
                                 ncol=1,rel_heights = c(0.5,1,1,1,0.2))

```

## Now, the correlation on real data.
This is more involved, so just loading the data s it was generated in `/scratch16/abattle4/ashton/snp_networks/presentation_figures/cohort_overlap_ukbb_may2024.R`
**NO LONGER INCLUDING THESE in the figs**
```{r, eval = FALSE}
load("/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/ukbb_overlap_dat.RData")
overlap.no.herit.half <- overlap.no.herit
overlap.no.herit.half[upper.tri(overlap.no.herit.half)] <- NA
noherit.corr <- plotCorrelationHeatmap(no.herit.query.corr, drop.labels=TRUE, col_limit=c(-1,1.2), font_size = 8,show.nums = TRUE,print.nums = overlap.no.herit.half) + 
  theme(legend.position = "right")+  labs(fill=bquote(g[cov_int]))  + theme(axis.text = element_blank()) + 
  annotate(geom="text",label = bquote(h^2 ~"= 0"),y = 12,x=4, size= 17/.pt)


overlap.selected.traits.half <- overlap.selected.traits
overlap.selected.traits.half[upper.tri(overlap.selected.traits.half)] <- NA
herit.corr <- plotCorrelationHeatmap(manual.query.corr, drop.labels=TRUE, col_limit=c(-1,1.2), font_size = 8,show.nums = TRUE,print.nums = overlap.selected.traits.half) + 
  theme(legend.position = "right")+  labs(fill=bquote(g[cov_int]))   + theme(axis.text = element_blank()) + 
  annotate(geom="text",label = bquote(h^2 ~"> 0"),y = 12,x=4, size= 17/.pt)

```
Okay, I need to merge the color schemes
Add in former figure 3 stuff:
```{r}
load("/scratch16/abattle4/ashton/snp_networks/presentation_figures/overlap_former_fig3.Rdata")
#bottom.plot <- cowplot::plot_grid(plotlist=list(tri.heatmap, orange.bars + theme(legend.position = "none")),labels = c("C",'D'),rel_widths = c(1,0.6))
bottom.plot <- cowplot::plot_grid(plotlist=list(tri.heatmap, orange.bars + theme(legend.position = "none")),rel_widths = c(1,0.6))
```


```{r}
library(ggpubr)
cor.shapes <- cowplot::plot_grid(blockheader, 
                                 oneb.plot, twob.plot, nob.plot, ggplot() + geom_blank() + theme_classic(),
                                 ncol=1,rel_heights = c(0.5,1,1,1,0.3))
#theme(plot.margin = unit(c(0,0,0,0), "cm"))
factor.sim <- cowplot::plot_grid(plotlist = list(arrows,u.sim ,cor.shapes), nrow=1, rel_widths = c(0.1,1.2,0.2),align = "v",axis = "tblr")
#upper.panel.corr <- cowplot::plot_grid(plotlist=list(noherit.corr+ theme(legend.position = "none"), herit.corr ),nrow  =1, rel_widths = c(0.8,1))
#upper.panel <- cowplot::plot_grid(plotlist = list(upper.panel.corr, gwas.based + theme(legend.position = "none")), rel_widths = c(1,1),labels = c('A', 'B'), rel_heights = c(0.6,1))
upper.panel <- cowplot::plot_grid(plotlist = list(gwas.based + theme(legend.position = "none")),bottom.legend, labels = c("",'A'),rel_widths = c(0.3,1))



cowplot::plot_grid(upper.panel, factor.sim, bottom.plot, rel_heights = c(0.5,1,0.6),nrow=3,labels = c("",'B',""))
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_JOINED.svg",
      width = 20,height=15,units="in")
#2000x1500
ggarrange(upper.panel, factor.sim,bottom.plot,
          labels = c("", "B",""),
          ncol = 1, nrow =3)

u.sim
```
Lazy approach- just save as images and import:
```{r}
upper.panel #1700 x 300
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_panelA.svg",
      width = 20,height=3.5,units="in")

u.sim #1300 x 450
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_panelB_1.svg",
      width = 11,height=4,units="in")
cor.shapes.noheader <- cowplot::plot_grid(oneb.plot, twob.plot, nob.plot,
                                 ncol=1,rel_heights = c(1,1,1,0.3)) #200 x 450
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_panelB_2.svg",
      width = 2,height=4,units="in")
cowplot::plot_grid(plotlist=list(tri.heatmap, orange.bars + theme(legend.position = "none"))) #
cowplot::plot_grid(plotlist = list(blank.arrows,u.sim ,cor.shapes), nrow=1, rel_widths = c(0.1,1.2,0.2),align = "v",axis = "tblr")
```




### Figures for poster:
Reduce to 3 panesl..
```{r}
for.poster <- factors.by.method.long.sub.plot %>% mutate("overlap_bins_3" = cut_interval(overlap_percent, n=3), "overlap_bins_2" = cut_interval(overlap_percent, n=2)) 
hum_names_tres <- as_labeller(
     c(`[0,33.3]` = "0-33%", `(33.3,66.7]` = "33-67%",`(66.7,100]` = "67-100%"))
hum_names_dos <- as_labeller(
     c(`[0,50]` = "0-50%", `(50,100]` = "50-100%"))

ggplot(for.poster, aes(x=method_label,y=as.numeric(value), fill=method_label))+
  geom_violin(aes(fill = method_label)) +theme_bw(17) +  geom_jitter(alpha = 0.1,aes(color=method_label))+
 fillScale + facet_wrap(~overlap_bins_2, nrow = 1,labeller = hum_names_dos, strip.position = "top") +  theme(strip.background =element_rect(fill="white")) + colScale + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),axis.title.x = element_blank()) + ylab(expression("# estimated factors" ~ hat(K))) + theme(legend.position = "none")
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_sim1_poster_overlap_version.svg",
      width = 8,height=4,units="in")

ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_sim1_poster_overlap_version.png",
      width = 8,height=4,units="in")
#800 x 400 I think?
```

Other sim fig- for talk?
```{r}
all.sims.sub.for.talk <- all.sims.joined.scaled %>% filter(N %in% c("1e+05", "10000"))%>% filter(Covar %in% c("none", "2b"))
ggplot(all.sims.sub.for.talk %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = yuan_U, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(16) +xlab("Factorization method")  + 
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + theme(legend.text=element_text(size=12)) + theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt"))
#950 x 550
```

For poster:

```{r}
all.sims.sub.for.poster <- all.sims.joined.scaled %>% filter(N %in% c("50000"))%>% filter(Covar %in% c("2b"))
ggplot(all.sims.sub.for.poster %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = yuan_U, fill = method_label)) +
  geom_boxplot()  +  theme_bw(20) +xlab("Factorization method")  + 
  theme(axis.text.x=element_text(angle=35, hjust=1, vjust=1, size=22)) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt")) 
#coord flip?
#830x550
twob.plot #
```


TODO: need to add in an additional entry for the missing SVD-elbow






```{r}
dd <- data.frame(x = c("A","B","C","D","E","F","G","H","I"),
                 y = c(1.782,0.136,0.978,0.645,0.518,1.474,0.855,0.673,0.369))
dd <- transform(dd, ylo = c(0.719,0.046,0.945,0.295,0.188,0.577,0.407,0.310,0.145), 
        yhi = c(4.420,0.398,1.012,1.411,1.424,3.768,1.798,1.460,0.940))

#Create a separate data frame for the arrow labels
arrowLab <- data.frame(lab = c("Increasing","Decreasing"),
                       x = c(0.15,0.15),y = c(10^0.25,10^(-0.25)))

ggplot(data = dd, aes(x=ordered(x, levels=rev(x)), y=y)) +
        geom_pointrange(aes(ymin=ylo, ymax=yhi)) + theme_classic() + geom_blank() +
        geom_segment(aes(x = 0, xend = 0, y= 1, yend= 2),
                         arrow=arrow(length=unit(0.2,"cm"))) +
        geom_segment(aes(x = 0, xend = 0, y= 1, yend= 10^(-0.25)),
                         arrow=arrow(length=unit(0.2,"cm"))) +
        geom_text(data = arrowLab,aes(x=x,y=y,label = lab),size = 3) 



```




