---
title: "Figure for cohort overlap and simulations."
output: html_notebook
---
Here, creating the simulation based figure
Note that previous sims were incorrect in that they didn't quite use the sd specifications in the right way. We should be able to get this to work
**Update the paths below to reflect the new locations, things have been mooved about!"**
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/simulation_processing_tools.R")
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/plot_functions.R")
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/factorization_methods.R")
source("/data/abattle4/aomdahl1/gene_utils/qqplot.R")
library(data.table)
library(magrittr)
library(dplyr)
library(stringr)
library(ggplot2)
library(RColorBrewer)
library(NatParksPalettes)
library(ggbeeswarm)
library(ggpubr)

```
## Null simulations of GWAS directly
Updated Oct 2024
```{r}
#this one was made in the old way, so just loaded the files directly in
load("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/null_matrix_spearman/N30000_multi-overlap_SD-1_h2-0-0_bmi-like.SCALED_Aug2024/null_covarfactorization.summaries.RData")
OLD_VERSION=FALSE
if(OLD_VERSION)
{
  #The format has changed a little- I had to do the whitened svd separately
load("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/null_matrix_spearman/N30000_multi-overlap_SD-1_h2-0-0_bmi-like.SCALED_Aug2024/null_covarfactorization.summaries.WHITENED_PCA.RData")
save.list$elbow.point <- elbow.point
save.list$flashr.runs <- flashr.runs
save.list$gd.point <- gd.point
save.list$GLEANR.runs.covar <- GLEANR.runs.covar
save.list$GLEANR.runs.std <- GLEANR.runs.std
save.list$sim.pca.alt <- sim.pca.alt
save.list$sim.pca.std <- sim.pca.std
save.list$kaiser.point <- kaiser.point
save.list.joined <- save.list
#rerun on flash:
}
save.list.joined <- save.list
load("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/null_matrix_spearman/N30000_multi-overlap_SD-1_h2-0-0_bmi-like.SCALED_Aug2024/null_covarfactorization.summaries.flash_SE_update.RData")
save.list.joined$flashr.runs.se <- save.list$flashr.runs.se
#temporary hack while bringing things up to the newest organization fo the software
tables.alt <- tableSimResultsFromList(save.list.joined)
#}else
#{
#  #save.list$flashr.runs <- save.list$flashr.runs.se
#  tables.alt <- tableSimResultsFromList(save.list)
#}

factors.by.method.long <- tables.alt$long
```
Now we can work as in the previous ones:
```{r}
#Do by percent of overlapping samples
factors.by.method.long %<>% 
  mutate("overlap_percent" = overlap_score/90000*100) %>%  
  mutate("overlap_bins" = cut_interval(overlap_percent, n=4)) %>% 
  filter(method != "whitened_gd.point", method != "gd.point")

```
Build specific color pallette
Choices here- dropping SVD-beta
```{r}
map.to.new <- data.frame(
  "method"=unique(factors.by.method.long$method),
"method_label"=c("GLEANR", "GLEANR-U", "SVD-adj (elbow)", 
               "SVD-adj (avg)","flash", "SVD (elbow)", "SVD (avg)"))

color.map.pub <-buildPubColors()
```

Map
```{r}
factors.by.method.long.sub.plot <- left_join(factors.by.method.long, map.to.new, by="method") %>% left_join(., color.map.pub, by="method_label")%>%  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% arrange(method_label)

sub.pub <- color.map.pub %>% mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% filter(method_label %in% factors.by.method.long.sub.plot$method_label)

```
Plot
```{r}

myColors <- color.map.pub$color
names(myColors) <- levels(factors.by.method.long.sub.plot$method_label)
colScale <- scale_colour_manual(name = "Method",values = myColors)
fillScale <- scale_fill_manual(name = "Method",values = myColors)


hum_names <- as_labeller(
     c(`[0,25]` = "0-25%", `(25,50]` = "25-50%",`(50,75]` = "50-75%", 
       `(75,100]` = "75-100%"))

gwas.based <- ggplot(factors.by.method.long.sub.plot, aes(x=method_label,y=as.numeric(value), fill=method_label))+
  geom_violin(aes(fill = method_label)) +theme_bw(15) +  geom_jitter(alpha = 0.1,aes(color=method_label))+
 fillScale + facet_wrap(~overlap_bins, nrow = 1,labeller = hum_names, strip.position = "top") +  theme(strip.background =element_rect(fill="white")) + colScale + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),axis.title.x = element_blank()) + ylab(expression("# estimated factors" ~ hat(K))) 
```
Great, Im' actually pretty happy with that.
##################################

## Factor-based simulations
Load in the results
```{r}
no.covar <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/no_overlap/")  %>% 
  mutate( "Covar"=  "none")

oneb.covar <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/1b_overlap/")  %>% 
  mutate("Covar"=  "1b")
twob.covar <- mergeTabularResults(query="",base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/2b_overlap/")  %>% 
  mutate("Covar"=  "2b")

all.sims.joined.scaled <- rbind(no.covar,oneb.covar,twob.covar)

```

Add in some fake rows so I can get the color legend for all methods

```{r}
all.sims.joined.scaled$N <- factor(all.sims.joined.scaled$N, levels = c("5000","10000","50000","1e+05","2e+05","mixed"))

full.list=c("SVD", "SVD (avg)", "SVD (elbow)","SVD-adj", "SVD-adj (avg)","SVD-adj (elbow)",
            "flash","flash (Z)","flash (Kronecker)","FactorGo","GLEANR-U","GLEANR")

drop.methods <- c()

#add fakes to make sure all methods are included.....
fake.row.svd.elbow <- all.sims.joined.scaled[1,]; fake.row.svd.elbow$method="SVD (elbow)";  fake.row.svd.elbow[,8:35] <- NA
fake.row.svd.adjelbow <- all.sims.joined.scaled[1,]; fake.row.svd.adjelbow$method="SVD-adj (elbow)";  fake.row.svd.adjelbow[,8:35] <- NA
fake.row.svd.adj_avg <- all.sims.joined.scaled[1,]; fake.row.svd.adj_avg$method="SVD-adj (avg)";  fake.row.svd.adj_avg[,8:35] <- NA
all.sims.joined.scaled <- rbind(all.sims.joined.scaled, fake.row.svd.elbow,fake.row.svd.adjelbow,fake.row.svd.adj_avg)
#all.sims.joined.unscaled <- rbind(all.sims.joined.unscaled, fake.row)

name.map = data.frame("method_label"=c("SVD", "flash (Z)", "flash", "flash (Kronecker)","SVD (avg)", "GLEANR", "GLEANR-U", "SVD-adj", "SVD (beta)", "FactorGo", "SVD (elbow)","SVD-adj (elbow)","SVD-adj (avg)"), "method" = unique(all.sims.joined.scaled$method))

level_order=name.map$method_label

all.sims.joined.scaled %<>% left_join(.,name.map, by="method") %>% 
  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% arrange(method_label) %>% filter(method != "SVD_beta",method_label != "flash (Z)")

```
Numerical checks- where do we outperform other methods?
```{r}
#Get for U- my claim holds true!
#need to update this- group by N and covar at the end.
all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianU) %>% filter(Covar == "1b") %>% group_by(N) %>% slice_max(medianU)

all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianU) %>% filter(Covar == "2b") %>% group_by(N) %>% slice_max(medianU)

all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianU) %>% filter(Covar == "none") %>% group_by(N) %>% slice_max(medianU)
```

Claim for V?
```{r}
#Get for V- my claim holds true!
all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianV) %>% filter(Covar == "1b") %>% group_by(N) %>% slice_max(medianV)

all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianV) %>% filter(Covar == "2b") %>% group_by(N) %>% slice_max(medianV)

all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)) %>% group_by(method_label, N, Covar) %>% summarize("medianV" = median(yuan_V), "medianU"=median(yuan_U)) %>% arrange(N, -medianV) %>% filter(Covar == "none") %>% group_by(N) %>% slice_max(medianV)
```

### Make the nice plots
The figure for V will go into supplement; figure for U in main text.
```{r}
##Labeller with updated name (1/31)
pop_names <- as_labeller(
     c(`5000` = "N<sub>S</sub>=5k", `10000` = "N<sub>S</sub>=10k",`50000` = "N<sub>S</sub>=50k", 
       `1e+05` = "N<sub>S</sub>=100k",`2e+05` = "N<sub>S</sub>=200k", "mixed"="N<sub>S</sub>=[10K,200K]",
     `1b`="1-block", `2b`="2-block", `none`="No overlap"))

# Supplemental figure
drop.methods <- c("flash (Z)", "SVD (elbow)","SVD (elbow)","SVD-adj (elbow)","SVD-adj (avg)", "SVD (avg)")

v.sim <- ggplot(all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)), aes(x = method_label, y = yuan_V, fill = method_label)) + geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(R^2 ~ "with simulated V")) + fillScale +  
  theme(strip.background =element_rect(fill="white"),legend.position = "top",strip.text = ggtext::element_markdown()) + ggtitle("Performance across simulations with respect to V")
v.leg <- ggpubr::get_legend(v.sim)

#Chat GPT helped me with this one.
u.sim.for.legend <- ggplot(all.sims.joined.scaled %>% filter(method_label != "flash (Z)"), 
                aes(x = method_label, y = yuan_U, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw() +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left",strip.text = ggtext::element_markdown()) +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + theme(legend.text=element_text(size=12))
#bottom.legend = get_plot_component(plot, 'guide-box-left', return_all = TRUE)
bottom.legend <- get_legend(u.sim.for.legend)
drop.methods <- c("flash (Z)", "SVD (elbow)","SVD (elbow)","SVD-adj (elbow)","SVD-adj (avg)", "SVD (avg)")
u.sim <- ggplot(all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = yuan_U, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left",strip.text = ggtext::element_markdown()) +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + 
  theme(legend.text=element_text(size=12),legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt")) + ggtitle("Covariance adjustment improves reconstruction of simulated genetic factors") + theme(plot.title = element_text(size=10,color="dimgrey"))  
```

Previous versions included arrows in plot- see version control for those. Have since been removed.

Results basically look the same, kinda. Scaled is more favorable for us, and I think its more fair all around.
#### Supplemenetal figure- looking at AUC, Kappa, F1:
Not doing Kappa or F1 here- Kappa just looks at agreement, when no known truth. F1 we don't have the precisoin/recall numbers close at hand.
```{r}
kappa.v  <- ggplot(all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = joint_cb_kappa_v, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(kappa ~ "for simulated V")) +
  fillScale +
  theme(strip.background =element_rect(fill="white"), legend.position = "left",strip.text = ggtext::element_markdown()) +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + 
  theme(legend.text=element_text(size=12)) + theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt"))  + 
  ggtitle("Classification (0/+/-) across simulations with respect to V")

kappa.u <- ggplot(all.sims.joined.scaled %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = joint_cb_kappa_u, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(kappa ~ "for simulated U")) +
  fillScale + 
  theme(strip.background =element_rect(fill="white"), legend.position = "left",strip.text = ggtext::element_markdown()) +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + 
  theme(legend.text=element_text(size=12),legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt"))  + 
  ggtitle("Classification (0/+/-) across simulations with respect to U")

cowplot::plot_grid(v.sim + theme(legend.position = "none"),kappa.v, kappa.u,v.leg,ncol=1,rel_heights = c(1,1,1,0.3)) 
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/suppl6_alternate_sim_metrics.svg",
      width = 8,height=12,units="in")
```



### Heatmaps of block structure
These will be placed to to the right of boxplots in Inkscape
```{r}
oneb.version <- fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/1b_overlap/V101_U101_MAF-mix_eur_N-10000_RHO-1b_high_mixed_p_No-1b_high_no/sim1.c_matrix.txt")

oneb.plot <- plotCorrelationHeatmap(as.matrix(oneb.version),typin = "none") + theme_minimal()+
  theme(axis.text.x=element_blank(),axis.text.y = element_blank(), legend.position = "none",
        axis.ticks = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=3)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(plot.margin = unit(c(-0.5,-1,-0.5,-1), "cm")) + coord_fixed()
  
```

```{r}
twob.version <- fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/2b_overlap/V101_U101_MAF-mix_eur_N-10000_RHO-2b_mid_mixed_p_No-2b_high_no/sim1.c_matrix.txt")

twob.plot <- plotCorrelationHeatmap(as.matrix(twob.version),typin = "none") + theme_minimal()+
  theme(axis.text.x=element_blank(),axis.text.y = element_blank(), legend.position = "none",
        axis.ticks = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=3)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(plot.margin = unit(c(-0.5,-1,-0.5,-1), "cm"))+ coord_fixed()
```

```{r}
nob.version <- fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/no_overlap/V101_U101_MAF-mix_eur_N-10000_RHO-none_No-none/sim1.c_matrix.txt")

nob.plot <- plotCorrelationHeatmap(as.matrix(nob.version),typin = "none") + theme_minimal()+
  theme(axis.text.x=element_blank(),axis.text.y = element_blank(), legend.position = "none",
        axis.ticks = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=3)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ theme(plot.margin = unit(c(-0.5,-1,-0.5,-1), "cm")) + coord_fixed()
```

```{r}
blockheader <- ggdraw() +
  draw_label("C in simulation",size = 9,color = "dimgrey") + theme(plot.margin = unit(c(0,0,-1,0), "cm"))

cowplot::plot_grid(blockheader, 
                                 oneb.plot, twob.plot, nob.plot, ggplot() + geom_blank() + theme_classic(),
                                 ncol=1,rel_heights = c(0.5,1,1,1,0.2))

```

## Now, the correlation on real data.
This is more involved, so just loading the data s it was generated in `/scratch16/abattle4/ashton/snp_networks/presentation_figures/cohort_overlap_ukbb_may2024.R`

Okay, I need to merge the color schemes
Add in former figure 3 stuff:
```{r}
load("/scratch16/abattle4/ashton/snp_networks/presentation_figures/overlap_former_fig3.Rdata")
#bottom.plot <- cowplot::plot_grid(plotlist=list(tri.heatmap, orange.bars + theme(legend.position = "none")),labels = c("C",'D'),rel_widths = c(1,0.6))
bottom.plot <- cowplot::plot_grid(plotlist=list(tri.heatmap, orange.bars + theme(legend.position = "none")),rel_widths = c(1,0.6))
```

Previously had tried getting everything programatically on one plot, which proved more trouble than it was worth. See version control for older versions.
Lazy approach- just save as images and import:
```{r}
upper.panel #1700 x 300
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_panelA.svg",
      width = 20,height=3.5,units="in")

u.sim #1300 x 450
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_panelB_1.svg",
      width = 11,height=4,units="in")
cor.shapes.noheader <- cowplot::plot_grid(oneb.plot, twob.plot, nob.plot,
                                 ncol=1,rel_heights = c(1,1,1,0.3)) #200 x 450
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_panelB_2.svg",
      width = 2,height=4,units="in")
cowplot::plot_grid(plotlist=list(tri.heatmap, orange.bars + theme(legend.position = "none"))) #
cowplot::plot_grid(plotlist = list(u.sim ,cor.shapes), nrow=1, rel_widths = c(0.1,1.2,0.2),align = "v",axis = "tblr")
```

### Figures for poster:
Reduce to 3 panesl..
```{r}
for.poster <- factors.by.method.long.sub.plot %>% mutate("overlap_bins_3" = cut_interval(overlap_percent, n=3), "overlap_bins_2" = cut_interval(overlap_percent, n=2)) 
hum_names_tres <- as_labeller(
     c(`[0,33.3]` = "0-33%", `(33.3,66.7]` = "33-67%",`(66.7,100]` = "67-100%"))
hum_names_dos <- as_labeller(
     c(`[0,50]` = "0-50%", `(50,100]` = "50-100%"))

ggplot(for.poster, aes(x=method_label,y=as.numeric(value), fill=method_label))+
  geom_violin(aes(fill = method_label)) +theme_bw(17) +  geom_jitter(alpha = 0.1,aes(color=method_label))+
 fillScale + facet_wrap(~overlap_bins_2, nrow = 1,labeller = hum_names_dos, strip.position = "top") +  theme(strip.background =element_rect(fill="white")) + colScale + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),axis.title.x = element_blank()) + ylab(expression("# estimated factors" ~ hat(K))) + theme(legend.position = "none")
ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_sim1_poster_overlap_version.svg",
      width = 8,height=4,units="in")

ggsave(filename="/scratch16/abattle4/ashton/snp_networks/presentation_figures/gleaner_manuscript_draft/figures/fig2_sim1_poster_overlap_version.png",
      width = 8,height=4,units="in")
#800 x 400 I think?
```

Other sim fig- for talk?
```{r}
all.sims.sub.for.talk <- all.sims.joined.scaled %>% filter(N %in% c("1e+05", "10000"))%>% filter(Covar %in% c("none", "2b"))
ggplot(all.sims.sub.for.talk %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = yuan_U, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(16) +xlab("Factorization method")  + 
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left",strip.text = ggtext::element_markdown()) +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) +
  theme(legend.text=element_text(size=12),legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt"))
#950 x 550
```

For poster:

```{r}
all.sims.sub.for.poster <- all.sims.joined.scaled %>% filter(N %in% c("50000"))%>% filter(Covar %in% c("2b"))
ggplot(all.sims.sub.for.poster %>% filter((!method_label %in% drop.methods)), 
                aes(x = method_label, y = yuan_U, fill = method_label)) +
  geom_boxplot()  +  theme_bw(20) +xlab("Factorization method")  + 
  theme(axis.text.x=element_text(angle=35, hjust=1, vjust=1, size=22)) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt")) 
#coord flip?
#830x550
twob.plot #
```

