---
title: "R Notebook"
output: html_notebook
---

```{r, setup}
pacman::p_load(dplyr, tidyr, data.table, magrittr, ggplot2, data.table)
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/plot_functions.R")
```
First- spot. check the analysis and matrix to see how well we filled it... if the new approach worked.
```{r}
rg.mat <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/summary_data/rg.tab.NO_SE_FILTERs.csv")
plotCorrelationHeatmap(as.matrix(rg.mat))
```
Visualize:
```{r}
trait.cats <- fread("/scratch16/abattle4/ashton/snp_networks/scratch/panUKBB_analysis/panUKBB_traits_manifest_v2.tsv")
trait.doc.run <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/trait_selections/panUKBB_complete.studies.tsv")

all(trait.doc.run$V2 %in% trait.cats$`Clean name`)
by.group.df <- trait.cats %>% filter(`Clean name` %in% trait.doc.run$V2) %>% group_by(Category) %>% summarize("num" = n())
ggplot(by.group.df, aes(x = reorder(Category,num), y= num)) + geom_bar(stat="identity", fill = "coral") + theme_bw(15) + 
  coord_flip() + ggtitle("Traits for consideration") + xlab("Trait categories") + ylab("# studies")
```


Spot check on numbers ?


Now, let's look at the problematic pairs
```{r}
rg.mat.test <- as.matrix(rg.mat)
diag(rg.mat.test) <- 0
keep.entries <- which(apply(rg.mat.test, 1, function(x) any(abs(x) > 0.8)))

plotCorrelationHeatmap(rg.mat.test[keep.entries, keep.entries])
```
Okay- as before, we are picking the ones we want:
```{r}
phenotype.h2 <- fread("/data/abattle4/lab_data/GWAS_summary_statistics/PanUKBB/Pan-UK_Biobank_phenotype_manifest-h2_manifest.tsv",sep = "\t") %>%
  filter(pop == "EUR")

map.traits <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/trait_selections/panUKBB_complete.studies.tsv")
map.traits$query = gsub(pattern = "/data/abattle4/lab_data/GWAS_summary_statistics/PanUKBB/phenotype_flat_files/EUR_only/", replacement = "", x=map.traits$V1) %>% gsub(pattern = ".tsv.EUR_ONLY.gz", x=., replacement = "")

phenotype.h2 <- phenotype.h2 %>% mutate("query" = 
                                          paste0(trait_type, "-", phenocode, "-", pheno_sex, 
                                                 ifelse(coding != "" , paste0("-", coding), ifelse(modifier != "", paste0("-", modifier), ""))))
stopifnot(all(map.traits$query %in% phenotype.h2$query))
h2.for.selected <- left_join(map.traits, phenotype.h2, by = "query") %>% filter(V2 %in% colnames(rg.mat)) %>% arrange(-estimates.final.h2_liability) %>% select(V2, V4,query,estimates.final.h2_z,estimates.final.h2_observed,estimates.final.h2_liability)


phenotype.filtered.pre <- fread("/data/abattle4/lab_data/GWAS_summary_statistics/PanUKBB/Pan-UK_Biobank_phenotype_manifest-phenotype_manifest.tsv", sep = "\t") %>% filter(phenotype_qc_EUR %in% c("PASS", "not_EUR_plus_1")) %>% 
  rowwise() %>%  mutate("Neff" = ifelse(is.na(n_controls_EUR), n_cases_EUR, psych::harmonic.mean(c(n_cases_EUR,n_controls_EUR)))) %>% ungroup() %>% 
  filter(Neff > 5000,pheno_sex== "both_sexes") %>% mutate("query" = 
                                          paste0(trait_type, "-", phenocode, "-", pheno_sex, 
                                                 ifelse(coding != "" , paste0("-", coding), ifelse(modifier != "", paste0("-", modifier), "")))) %>%
  select(query, Neff)
#Pick out just the higher NEFF ones, as we specified earlier
h2.for.selected <- left_join(h2.for.selected,phenotype.filtered.pre,by="query") %>% group_by(V2) %>% slice_max(Neff) %>% ungroup() %>% arrange(-estimates.final.h2_liability)
```

Now, we repeat the procedure from previously. Include p-values.
```{r}
rg.p.values <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/summary_data/rg_pvals.csv")
stopifnot(all(colnames(rg.p.values) == colnames(rg.mat)))
rg.mat <- as.matrix(rg.mat);rg.p.values <- as.matrix(rg.p.values)
```
Look at high correlation rg with low p-values?
```{r}

plot(as.vector(rg.mat), -log10(as.vector(rg.p.values)), xlab = bquote(r[g]), ylab = "-log10(p)"); abline(h = -log10(0.01), col = "blue")
#rg.mat[rg.p.values > 0.01] <- 0
```
Now, by heritability
```{r}
head(h2.for.selected)
rownames(rg.mat) <- colnames(rg.mat)
rownames(rg.p.values) <- colnames(rg.p.values)
rg.mat.long <- reshape2::melt(rg.mat) %>% set_colnames(c("p1", "p2", "rg"))
p.mat.long <- reshape2::melt(rg.p.values) %>% set_colnames(c("p1", "p2", "p"))
if(all(rg.mat.long$p1 == p.mat.long$p1) & all(rg.mat.long$p2 == p.mat.long$p2))
   {
     rg.mat.long$p <- p.mat.long$p
}

#porting the code over from the previous
buildTraitSetFinal <- function(pheno.data, corr.data, thresh)
{
  #Use the other version, gives full report:
  source("/scratch16/abattle4/ashton/snp_networks/scratch/panUKBB_analysis/cor_matching_function_helper.R")
  match.set <- c()
for(i in 1:nrow(pheno.data))
{
  #if(i %% 100 == 0) print(i)
  get_id <- pheno.data[i,]$V2
  if(passCorCheck(get_id, match.set, corr.data, thresh = thresh))
  {
    match.set <- c(match.set, get_id)
  }
}
match.set
}

```
Now pick out the ones...


```{r}
sink('/scratch16/abattle4/ashton/snp_networks/scratch/panUKBB_analysis/part2_filtering_reasons_0.8.txt', append=FALSE)
inclusive.8 <- buildTraitSetFinal(h2.for.selected,rg.mat.long,0.8)
sink()
sink('/scratch16/abattle4/ashton/snp_networks/scratch/panUKBB_analysis/part2_filtering_reasons_0.7.txt', append=FALSE)
inclusive.7 <- buildTraitSetFinal(h2.for.selected,rg.mat.long,0.7)
sink()
```
Visualize these in a barplot:
```{r}
by.group.df <- trait.cats %>% filter(`Clean name` %in%inclusive.7) %>% group_by(Category) %>% summarize("num" = n())
ggplot(by.group.df, aes(x = reorder(Category,num), y= num)) + geom_bar(stat="identity", fill = "skyblue") + 
  theme_bw(15) + coord_flip() + ggtitle("Traits selected for GLEANER") + ylab("# studies") + xlab("Trait categories")
```
For alexis:
```{r}
for.alexis <- left_join(h2.for.selected %>% rename("clean_name"=V2), trait.cats %>% filter(`Clean name` %in% trait.doc.run$V2) %>% rename("clean_name"=`Clean name`), by = "clean_name") %>% select(clean_name, Category, sample_size, description, description_more,estimates.final.h2_liability,estimates.final.h2_z, Neff ) %>%
  mutate("To_evaluate_in_GLEANER"=ifelse(clean_name %in% inclusive.7, TRUE, FALSE))
View(for.alexis)
write.table(for.alexis, file = "/scratch16/abattle4/ashton/snp_networks/scratch/panUKBB_analysis/traits_for_review.txt", quote = FALSE, row.names = FALSE, sep = "#")
```

## Overlap effect
hat does the overlap effect look like on those guys?
```{r}
gcov_fdr <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/summary_data/gcov_int.tab.csv"))
gcov_se <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/summary_data/gcov_int_se.tab.csv"))
plotCorrelationHeatmap(as.matrix(gcov_fdr))
```
Subset to the 0.7 cutoff list:
```{r}
i.c <- which(colnames(gcov_fdr) %in% inclusive.7)
plotCorrelationHeatmap(as.matrix(gcov_fdr)[i.c,i.c],drop.labels = TRUE) + theme_bw(16) + scale_y_discrete(guide = guide_axis(n.dodge = 2)) + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
rownames(gcov_fdr) <- colnames(gcov_fdr)
rownames(gcov_se) <- colnames(gcov_se)
gcov_fdr_7 <- gcov_fdr[inclusive.7,inclusive.7]
gcov_se_7 <- gcov_se[inclusive.7,inclusive.7]
```
So some of these have really exceptionally high overlap effects... what do I do with those?
```{r}
gcov_fdr_7[which(gcov_fdr_7 > 1)] #hmm some of these estimates are problematic. Like they are way above 1.
gcov_se_7[which(gcov_fdr_7 > 1)] #A few of these are too high...
exceeds <- apply(gcov_fdr_7,2,function(x) which(abs(x) > 1))
danger.names <- c()
for(i in 1:length(exceeds))
{
  pheno = exceeds[[i]]
  n = names(exceeds)[i]
  if(length(pheno) > 0)
  {
    print(paste(n, pheno))
    danger.names <- c(danger.names,n)
  }
}
plotCorrelationHeatmap(gcov_fdr_7[danger.names, danger.names])
plotCorrelationHeatmap(rg.mat[danger.names, danger.names])
```



3 Phenotypes had estimated cohort overlap inflation that was very high, over 1.

How to deal with this? We can assume they have complete cohort overlap here (Ns = N1=N2), but it seems unlikely that the phenotype correlation is THAT HIGH.
Options- ignore, adjust to 1 or 0.99, or lwoer threhsold...
```{r}
h2.for.selected
```

I think we have an issue with some genetic correlation estimates being low. Like these traits should be much more highly correlated, no?
#Load in all the data, and see what the z-scores look like....
```{r}
file.list <- paste0("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/rg_ldsr/tabular/",
                    list.files(path = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/rg_ldsr/tabular/", pattern = "*.rg_report.tsv"))
rg.all <- do.call("rbind", lapply(file.list, fread))

fixStrings <- function(x)
{
  
  #splitted <- str_split(x, pattern = "_")[[1]]
   #paste(unlist(splitted[2:length(splitted)]), collapse="_")
   splitted <- str_split(x, pattern = "\\.")[[1]]
   splitted[4]
  
}

rg.all$p1 <- gsub(x=rg.all$p1, pattern = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/panUKBB_complete/", replacement = "") %>% sapply(., fixStrings)
  #gsub(x=., pattern = "both_sexes-irnt.tsv.EUR_ONLY.", replacement = "") %>% gsub(x=., pattern = ".sumstats.gz", replacement = "") %>% sapply(., fixStrings)

rg.all$p2 <- gsub(x=rg.all$p2, pattern = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/panUKBB_complete/", replacement = "") %>% sapply(., fixStrings)
  #gsub(x=., pattern = "both_sexes-irnt.tsv.EUR_ONLY.", replacement = "") %>% gsub(x=., pattern = ".sumstats.gz", replacement = "") %>% sapply(., fixStrings)

View(rg.all)
```
Now just look at those in our list
```{r}
View(rg.all %>% filter(p1 %in% inclusive.7, p2 %in% inclusive.7))
write.table(x= data.frame("pheno_keep"=inclusive.7), row.names = FALSE, col.names = FALSE, quote = FALSE, file = "/scratch16/abattle4/ashton/snp_networks/scratch/panUKBB_analysis/rg_filtered_0.7_traits.txt")

```

FINAL VISUALS:
```{r}
rg.mat <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/summary_data/rg.tab.NO_SE_FILTERs.csv"))
rownames(rg.mat) <- colnames(rg.mat)
plotCorrelationHeatmap(rg.mat[inclusive.7, inclusive.7]) + theme_bw(16) + scale_y_discrete(guide = guide_axis(n.dodge = 2))

head(phenotype.h2)

ggplot(for.alexis, aes(x = reorder(clean_name,as.numeric(estimates.final.h2_liability)), y = estimates.final.h2_liability)) + geom_bar(stat = "identity", fill = "coral") +  
  ylab(bquote("Liability scale" ~h^2 )) + xlab("GWAS study") + theme_bw(16) + scale_x_discrete(guide = guide_axis(n.dodge = 2)) + coord_flip()
mean(for.alexis$estimates.final.h2_liability)

summary(h2.for.selected$estimates.final.h2_liability)
#+ theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))
```

## Manually get the new set of files
First, the missingness report
```{r}
mreport <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/panUKBB_complete/missingness_report.COMPLETE.tsv")
mreport$names <- sapply(str_split(mreport$V1, pattern ="\\."), function(x) x[length(x)-2])
write.table(filter(mreport, names %in% inclusive.7) %>% select(V1, V2), file = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/panUKBB_complete/missingness_report.tsv", row.names = FALSE, col.names = FALSE,quote = FALSE)
```
Next, the SD report
```{r}
sdreport <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/panUKBB_complete/sample_sd_report.tsv")
sdreport$names <- sapply(str_split(sdreport$V1, pattern ="\\."), function(x) x[length(x)-2])
file.rename("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/panUKBB_complete/sample_sd_report.tsv", "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/panUKBB_complete/sample_sd_report.OLD.tsv")

write.table(filter(sdreport, names %in% inclusive.7) %>% select(V1, V2), file = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/panUKBB_complete/missingness_report.tsv", row.names = FALSE, col.names = FALSE,quote = FALSE)
```


Next, the gcov matrix
```{r}
write.table(data.frame((gcov_fdr[inclusive.7,inclusive.7])) %>% set_colnames(inclusive.7), row.names = FALSE, quote = FALSE, file = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/summary_data/gcov_int.tab.SELECTED.csv")

write.table(data.frame((gcov_se[inclusive.7,inclusive.7])) %>% set_colnames(inclusive.7), row.names = FALSE, quote = FALSE, file = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/ldsr_results/panUKBB_complete/summary_data/gcov_int_se.tab.SELECTED.csv")
```


## Review of the SNPs selected:
```{r}

```

