---
title: "Comparing (RAW) SNP data, v2"
output: html_notebook
---

This notebook is the v2 analog of "comparing_raw_data.Rmd", with all the benefits of hindsight.
Phenotypes have been selected in a much more stringent manner, so comparison of LDSC results is not done here (See the results in local notebookk, "ukbb_finngen_phenotypes.Rmd").
The main aim here is to determine which subset of SNPs to use, and if my current rhesholds are sufficient.

```{r}
source("~/scratch16-abattle4/ashton/snp_networks/custom_l1_factorization/src/plot_functions.R")
source("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/src/evaluateSimR2.R")
pacman::p_load(ggplot2, magrittr, tidyr)
```

Read in the SNPs, pre pruning.
Note, we need to do the inngen_benchmark.union. VERSION TO KEEP IT IN RSIDS, THE SNP IDS WON'T work for the previous lookups.

```{r}
#snps.finngen <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/finngen_benchmark_2/finngen_benchmark_2.1000G.txt", header = F)$V1
#snps.ukbb <-  fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/ukbb_benchmark_2.1000G.txt", header = F)$V1

snps.finngen <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/finngen_benchmark_2/p1e-5/finngen_benchmark_2.union.txt", header = F)$V1
snps.ukbb <-  fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/p1e-5/ukbb_benchmark_2.union.txt", header = F)$V1

length(snps.finngen)
length(snps.ukbb)
```

Unsurprisingly, FG has alot more SNPs `(51755-34518)/34518= 49.9%` more to be exact

What is the overlap between these?
```{r}
joined.list <-list("Finngen"=snps.finngen, "UKBB"=snps.ukbb)
snps.in.both <- snps.finngen[snps.finngen %in% snps.ukbb]
snps.in.fg.only <- snps.finngen[!(snps.finngen %in% snps.ukbb)]
snps.in.uk.only <- snps.ukbb[!(snps.ukbb %in% snps.finngen)]
length(union(snps.finngen, snps.ukbb))
library(ggvenn)
ggvenn(
  joined.list, 
  fill_color = c("#0073C2FF", "#EFC000FF"),
  stroke_size = 0.5
  )
```
Thinking about why this is-
- Some small number not sampled in both (based on previous work, not many) (somewhere between)


How about the pruned?
```{r}
snps.finngen.pruned <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/finngen_benchmark_2/p1e-5/finngen_benchmark_2.250kb.0.2r2.prune.in", header = F)$V1
snps.ukbb.pruned <-  fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/p1e-5/ukbb_benchmark_2.250kb.0.2r2.prune.in", header = F)$V1
length(snps.finngen.pruned)
length(snps.ukbb.pruned)
```

Once again, FG has alot more SNPs `(6864-4004)/4004= 71.4%` more to be exact

What is the overlap between these?
```{r}
joined.list.pruned <-list("Finngen"=snps.finngen.pruned, "UKBB"=snps.ukbb.pruned)
library(ggvenn)
ggvenn(
  joined.list.pruned, 
  fill_color = c("#0073C2FF", "#EFC000FF"),
  stroke_size = 0.5
  )
```
I think what I'd like to do, to get us closer to like ~10,000 SNPs or so for the downstream analysis, be as inclusive as possible.
Options:
- Drop threshold to 1e-4 and re-run (probably the easiest)
- Keep any SNPs in one that aren't in the other that are at least 1e-4 in the other (a bit more involved)
   - how could I do this quickly? Run an p-value extract on all the SNPs that are in the other and use that one.
   Let's try it.
```{r}
finngen.exclusives <- data.frame("rsid" = snps.in.fg.only)
ukbb.exclusives <- data.frame("rsid" = snps.in.uk.only)

#Write these out and search for them
write.table(finngen.exclusives, col.names = FALSE, row.names = TRUE, quote = FALSE,
            file = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/finngen_benchmark_2/finngen_benchmark_2.p1e-5exclusives.txt")


write.table(ukbb.exclusives, col.names = FALSE, row.names = TRUE, quote = FALSE,
            file = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/ukbb_benchmark_2.p1e-5_exclusives.txt")
```
  
   
   don't run this in R please
```{bash, eval=FALSE}
ml anaconda
conda activate std
FG=finngen_benchmark_2
UK=ukbb_benchmark_2
python src/quickGWASIter.py  --type ldsc_custom --output gwas_extracts/${FG}/${UK}.p1e-5_exclusives \
--gwas_list gwas_extracts/${FG}/missingness_report.tsv --snp_list gwas_extracts/${UK}/${UK}.p1e-5_exclusives.txt \
--extension "sumstats.gz" --gwas_dir gwas_extracts/${FG}/

python src/quickGWASIter.py  --type ldsc_custom --output gwas_extracts/${UK}/${FG}.p1e-5_exclusives \
--gwas_list gwas_extracts/${UK}/missingness_report.tsv --snp_list gwas_extracts/${FG}/${FG}.p1e-5exclusives.txt \
--extension "sumstats.gz" --gwas_dir gwas_extracts/${UK}/
```

With this set, I intend to evaluate these SNPs against those selected from the UKBB/Finngen.
Look at the ones in A and not B- are the p-values close?
If they are close (i.e. 1e-4), keep them.
```{r}
original.finngen <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/finngen_benchmark_2/p1e-5/finngen_benchmark_2.p.tsv")
trait.names.ogf <- unlist(stringr::str_match(colnames(original.finngen)[-1], ".([\\w_]+).sumstats.gz")[,2])
#make sure the condition is right:
all(apply(original.finngen[,2:43], 1, function(x) { x[is.na(x)] <- 1; any(x < 1e-5)}))

omitted.from.ukbb <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/finngen_benchmark_2.p1e-5_exclusives.p.tsv")
trait.names.ofu <- unlist(stringr::str_match(colnames(omitted.from.ukbb)[-1], ".([\\w_]+).sumstats.gz")[,2])
all(trait.names.ofu == trait.names.ogf)
#First- make sure this omitted set lines up with the lists above
head(omitted.from.ukbb)
print(paste("Proportion included:", sum(finngen.exclusives$rsid %in% omitted.from.ukbb$ids)/length(omitted.from.ukbb$ids)))
```
Look at the distribution of p-values amongs those that don't match:
```{r}

sum(apply(omitted.from.ukbb[,2:43], 1, function(x) { x[is.na(x)] <- 1; min(x)}) < 1e-5)
sum(apply(omitted.from.ukbb[,2:43], 1, function(x) { x[is.na(x)] <- 1; min(x)}) < 1e-4)
sum(apply(omitted.from.ukbb[,2:43], 1, function(x) { x[is.na(x)] <- 1; min(x)}) < 1e-3)

dist.df <- data.frame(omitted.from.ukbb) %>% set_colnames(c("rsid",trait.names.ofu)) %>% pivot_longer(cols = trait.names.ofu)
ggplot(dist.df, aes(y = -log10(value), x= reorder(name, (value)))) + geom_boxplot() +theme_bw() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + geom_hline(yintercept = -log10(1e-5)) + ylab("-log10(pvalue)") + xlab("Study") + ggtitle("P-value distribution of UKBB SNPs that are p<1e-5 in Finngen but p>=1e-5 in UKBB")
``` 



Let's compare the p-values as a reference (across all SNPs)
```{r}

original.finngen.df <- data.frame(original.finngen) %>% filter(ids %in% finngen.exclusives$rsid) %>% 
  set_colnames(c("rsid",trait.names.ofu)) %>% mutate("Source" = "Finngen") %>% 
  pivot_longer(cols = all_of(trait.names.ofu), names_to = "study", values_to = "pval")%>% arrange(rsid, study)

omitted.from.ukbb.df <- data.frame(omitted.from.ukbb)  %>% set_colnames(c("rsid",trait.names.ofu)) %>%  dplyr::mutate("Source" = "UKBB") %>%
  pivot_longer(cols = all_of(trait.names.ofu), names_to = "study", values_to = "pval") %>% arrange(rsid, study)

#because the finngen ones are just the pruned ones, its not a complete sample
omitted.from.ukbb.df.sub <- omitted.from.ukbb.df %>% filter(rsid %in% unique(original.finngen.df$rsid)) %>% arrange(rsid, study)
stopifnot(all(original.finngen.df$rsid == omitted.from.ukbb.df.sub$rsid))
stopifnot(all(original.finngen.df$study == omitted.from.ukbb.df.sub$study))
plot(original.finngen.df$pval, omitted.from.ukbb.df.sub$pval)

finngen.exclusives.joined.df <- rbind(data.frame(original.finngen) %>% filter(ids %in% finngen.exclusives$rsid) %>% 
                                        set_colnames(c("rsid",trait.names.ofu)) %>% mutate("Source" = "Finngen"),
                   data.frame(omitted.from.ukbb)  %>% set_colnames(c("rsid",trait.names.ofu)) %>%  filter(rsid %in% unique(original.finngen.df$rsid)) %>% 
                     dplyr::mutate("Source" = "UKBB")) %>%
  pivot_longer(cols = all_of(trait.names.ofu), names_to = "study", values_to = "pval")
head(finngen.exclusives.joined.df)
ggplot(finngen.exclusives.joined.df, aes(x=study, y= -log10(pval), fill = Source)) + geom_boxplot() + ggtitle("Finngen SNPs are generally more significant\n(unsurprising, they were selected to be so)")
#head(finngen.exclusives.joined.df)
```
My question is- do the effects correlate? For a given study or a given SNP?
```{r}
original.finngen.wide <- data.frame(original.finngen) %>% filter(ids %in% finngen.exclusives$rsid) %>% 
  set_colnames(c("rsid",trait.names.ofu)) %>% mutate("Source" = "Finngen") %>% arrange(rsid)

omitted.from.ukbb.wide <- data.frame(omitted.from.ukbb)  %>% set_colnames(c("rsid",trait.names.ofu)) %>%  
  dplyr::mutate("Source" = "UKBB") %>% filter(rsid %in% unique(original.finngen.wide$rsid)) %>% arrange(rsid)

#Ensue our condition holds
sum(apply(omitted.from.ukbb.wide[,2:43], 1, function(x) { x[is.na(x)] <- 1; min(x)}) < 1e-5)
sum(apply(omitted.from.ukbb.wide[,2:43], 1, function(x) { x[is.na(x)] <- 1; min(x)}) < 1e-4)
sum(apply(omitted.from.ukbb.wide[,2:43], 1, function(x) { x[is.na(x)] <- 1; min(x)}) < 1e-3)
#revisit prior sanity checks... It all chc
hist(apply(omitted.from.ukbb.wide[,2:43], 1, function(x) { x[is.na(x)] <- 1; min(x)}),breaks = 50, xlab="pval")
#Make sure IDs line up.
stopifnot(all(omitted.from.ukbb.wide$rsid == original.finngen.wide$rsid))
head(original.finngen.wide)
head(omitted.from.ukbb.wide)
all.cors <- lapply(trait.names.ofu, function(x) cor.test(x=original.finngen.wide[,x],omitted.from.ukbb.wide[,x]))
cor.metric <- data.frame("study" = trait.names.ofu, 
           "pval" = sapply(all.cors, function(x) x$p.value),
           "cor" = sapply(all.cors, function(x) x$estimate),
           "upper" = sapply(all.cors, function(x) x$conf.int[2]),
           "lower" = sapply(all.cors, function(x) x$conf.int[1]))
ggplot(cor.metric, aes(x = reorder(study,cor), y = cor)) + geom_bar(stat = "identity") +  
  geom_errorbar( aes(ymin = lower, ymax = upper), data = cor.metric, width = 0.2) + theme_bw() + ggtitle("Correlation of p-values between (p<1e-5) Finngen\nand (unspecified) UKBB SNPs ")+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  ylab("Correlation of p-values") + xlab("Studies")
```
^Note- this is SNPs that aren't both passing the same significance threshold.
Can we look at that subset instead?
1) Look at just those SNPs passing the 1e-4 threshold in at least 1 study.
```{r}
head(omitted.from.ukbb.wide)
head(original.finngen.wide)
#Check assumptions abouf finngen
stopifnot(all(apply(original.finngen.wide[,2:43], 1, function(x) any(x < 1e-5)),na.rm = TRUE))
omitted.from.ukbb.wide
#Get rows passing the rule
keep.rows <- apply(omitted.from.ukbb.wide[,2:43], 1, function(x) { x[is.na(x)] <- 1; any(x < 1e-4)}) #this cannot be.
all(keep.rows, na.rm = TRUE)

which.correspond <- apply(omitted.from.ukbb.wide[,2:43], 1, function(x) { x[is.na(x)] <- 1; which(x < 1e-4)[1]})
contributing.phenos <- table(colnames(omitted.from.ukbb.wide[,2:44])[which.correspond])
contributing.phenos
dim(contributing.phenos)
```
All but 2 phenotypes are contributing to this.
I wonder if we should drop those 2?
```{r}
colnames(omitted.from.ukbb.wide[,2:43])[!(colnames(omitted.from.ukbb.wide[,2:43]) %in% names(contributing.phenos))]
```
Okay, might not be a bad idea to drop these 2, unless we see some correlation in these plots:
```{r}
keep.rsid<- omitted.from.ukbb.wide$rsid[apply(omitted.from.ukbb.wide[,2:43], 1, function(x) { x[is.na(x)] <- 1; any(x < 1e-4)})] #this cannot be.
original.finngen.wide.1e04 <- original.finngen.wide %>% filter(rsid %in% keep.rsid)
omitted.from.ukbb.wide.1e04 <- omitted.from.ukbb.wide %>% filter(rsid %in% keep.rsid)
getCrosswiseCors <- function(first, second,thresh,names)
{
  all.cors.1e04 <- lapply(names, function(x) cor.test(x=first[,x],second[,x]))
cor.metric.1e04  <- data.frame("study" = names, 
           "pval" = sapply(all.cors.1e04, function(x) x$p.value),
           "cor" = sapply(all.cors.1e04, function(x) x$estimate),
           "upper" = sapply(all.cors.1e04, function(x) x$conf.int[2]),
           "lower" = sapply(all.cors.1e04, function(x) x$conf.int[1]))
cor.metric.1e04
}
cor.metric.1e04_ <- getCrosswiseCors(original.finngen.wide.1e04,omitted.from.ukbb.wide.1e04, 1e-4,trait.names.ofu)

ggplot(cor.metric.1e04, aes(x = reorder(study,cor), y = cor)) + geom_bar(stat = "identity") +  
  geom_errorbar( aes(ymin = lower, ymax = upper), data = cor.metric.1e04, width = 0.2) + theme_bw() + ggtitle("Correlation of p-values between (p<1e-5) Finngen\nand (p<1e-4) UKBB SNPs ")+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  ylab("Correlation of p-values") + xlab("Studies")
```



Weird. So the subset of just pruned SNPs
2) Look at all the SNPs that do pass the thresholds in both
So it seems like this set of SNPs is really uninformative for traits on the far left, and pretty good for everything from colorectal cancer on.

Only 30 phenotypes contribute to this condition- if its 1e-5 in 1, its 1e-4 in other.
How many traits are contributing
```{r}
original.finngen.wide
which.correspond.fg <- apply(original.finngen.wide[,2:43], 1, function(x) { x[is.na(x)] <- 1; which(x < 1e-5)[1]})
by.name <- colnames(omitted.from.ukbb.wide[,2:44])[which.correspond]
```
What is our reference point?
```{r}
original.ukbb <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/ukbb_benchmark_2.p.tsv")
ukbb.in.fg <- data.frame(original.ukbb) %>% filter(ids %in% original.finngen$ids) %>% set_colnames(c("rsid",trait.names.ofu))
finngen.in.uk <- data.frame(original.finngen)  %>% filter(ids %in% ukbb.in.fg$rsid)%>% set_colnames(c("rsid",trait.names.ofu))
cross.1e5 <- getCrosswiseCors(ukbb.in.fg,finngen.in.uk, 1,trait.names.ofu)


ggplot(cross.1e5, aes(x = reorder(study,cor), y = cor)) + geom_bar(stat = "identity") +  
  geom_errorbar( aes(ymin = lower, ymax = upper), data = cross.1e5, width = 0.2) + theme_bw() + ggtitle("Correlation of p-values between (p<1e-5) Finngen\nand (p<1e-5) UKBB SNPs ")+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  ylab("Correlation of p-values") + xlab("Studies")
```


I wonder if I was to do a simple PCA
**Question**- if you wanted to make the analysis as similar as possible, how woudl you do it?
- Run PCA on each and drop out phenotypes that have very low reconstruction error.
- That might not be the mest bmetric- just might be low genetic signal, noisy, etc. Well that is what we are going foro, so maybe that's what to doo
Maybe we look at z-scores instead? later

After more than enough hemming and hawing, I think I will stick with the smaller set of SNPs as a test, to see what comes out of it.
If it goes well, great.
If not, circle back to THIS POINT and increase the # of SNPs.

Remove the HLA region
Get the joined list

Move on.
Now, just going to proceed with joint of the 2:
```{r}
snps.finngen <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/finngen_benchmark_2/finngen_benchmark_2.1000G.txt", header = F)$V1
snps.ukbb <-  fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/ukbb_benchmark_2.1000G.txt", header = F)$V1

length(snps.finngen)
length(snps.ukbb)
```

FG has `(47995-32172)/32172= 49.9%` more SNPs

What is the overlap between these?
```{r}
joined.list <-list("Finngen"=snps.finngen, "UKBB"=snps.ukbb)
snps.in.both <- snps.finngen[snps.finngen %in% snps.ukbb]
snps.in.fg.only <- snps.finngen[!(snps.finngen %in% snps.ukbb)]
snps.in.uk.only <- snps.ukbb[!(snps.ukbb %in% snps.finngen)]
length(union(snps.finngen, snps.ukbb))
library(ggvenn)
ggvenn(
  joined.list, 
  fill_color = c("#0073C2FF", "#EFC000FF"),
  stroke_size = 0.5
  )
write.table(data.table("rsids"=snps.in.both), file="/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/snps_in_both.txt", 
            row.names = FALSE, quote = FALSE, col.names = FALSE)

write.table(data.table("rsids"=snps.in.both), file="/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/finngen_benchmark_2/snps_in_both.txt", 
            row.names = FALSE, quote = FALSE, col.names = FALSE)
```
One last bar plot for visualization purposes- the correlation of the output snps....
```{r}
final.finngen <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/finngen_benchmark_2/finngen_benchmark_2_conservative.z.tsv")
trait.names.ofu <- unlist(stringr::str_match(colnames(final.finngen)[-1], ".([\\w_]+).sumstats.gz")[,2])
final.ukbb <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/gwas_extracts/ukbb_benchmark_2/ukbb_benchmark_2_conservative.z.tsv")
stopifnot(all(final.ukbb$ids==final.finngen$ids))
colnames(final.ukbb) <- c("rsid",trait.names.ofu)
colnames(final.finngen) <- c("rsid",trait.names.ofu)
cross.1e5 <- getCrosswiseCors(data.frame(final.ukbb),data.frame(final.finngen), 1,trait.names.ofu)


ggplot(cross.1e5, aes(x = reorder(study,cor), y = cor)) + geom_bar(stat = "identity") +  
  geom_errorbar( aes(ymin = lower, ymax = upper), data = cross.1e5, width = 0.2) + theme_bw() + ggtitle("Correlation of z-scores between intersecting Finngen\nand  UKBB pruned SNPs (p<1e-5) ")+ coord_flip() +
  ylab("Correlation of Z-scores") + xlab("Studies")

theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
```

