---
title: "L1_eigenfactors_weighted"
output: html_notebook
---
May 5, 2021
Here I'm developing my leading method idea for L1 ALS with first column based on eigenvalue, weighting based on e.
```{r}
pacman::p_load(data.table, tidyr, dplyr, readr, ggplot2, stringr, penalized, cowplot, parallel, doParallel)
source("~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/sn_spMF/fit_F.R")
source("~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/sn_spMF/update_FL_working.R")
source("~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/sn_spMF/fit_L.R")
source("/Users/ashton/Documents/JHU/Research/LocalData/snp_network/matrix_sparsification/plot_functions.R")
library(flashr)
source('~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/sn_spMF/compute_obj.R')
```

## Read in Data
we are not normalizing by any metric here...
This, of course, introduces a new set of pains. Becauuuuuseee the LDSC sum stats don't have the SE estimates, just the Z-scores. So we will need to pull this information out of the full summary stats.
Fortunately, I have scripts for this!
 For now, I want to leave local. But at some point, I want to move to MARCC
```{r}
z_scores <- fread("~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1_vars1e-5.FULL_LIST.1e-5.z.tsv") %>% drop_na() %>% arrange(ids)
names <- scan("~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1.names.tsv", what = character())
z <- as.matrix(z_scores %>% select(-ids))
any(duplicated(z_scores$ids))
```
Get the results locally, after running on MARCC
(for details on this, please see `may_2021.md` tracking document for 5/10)
```{bash}
source ~/.bash_profile
downloadFromBluecrab /work-zfs/abattle4/ashton/snp_networks/gwas_decomp_ldsc/gwas_extracts/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1_vars1e-5.FULL_LIST.1e-5.*.tsv ~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/
downloadFromBluecrab /work-zfs/abattle4/ashton/snp_networks/gwas_decomp_ldsc/gwas_extracts/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1_vars1e-5.pruned_rsids.txt  ~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/
ls ~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/
```
In this analysis, I wish to test scaling in 4 different ways:
 1. Betas alone
 2. Z-scores alone
 3. Betas with SE scaling
 4. Betas with var(MAF) scaling
Okay, now let's load the data.
Note that the SNPs differ here a bit from those in the z_scores
```{r}
rsid_map <- fread("~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1_vars1e-5.pruned_rsids.txt", header =FALSE) %>% rename("ids" = V1, "rsids" = V2)

W_se <- fread("~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1_vars1e-5.FULL_LIST.1e-5.se.tsv") %>% drop_na() %>% filter(ids %in% z_scores$ids) %>% arrange(ids) %>% select(-ids) %>% as.matrix()

W_maf <- fread("~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1_vars1e-5.FULL_LIST.1e-5.maf.tsv") %>% drop_na() %>% filter(ids %in% z_scores$ids) %>% arrange(ids) %>% select(-ids) %>% as.matrix()

W_varmaf <- matrix(apply(W_maf, 2, function(x) 2*x*(1-x)), nrow = nrow(W_maf), ncol = ncol(W_maf))

betas <- fread("~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1_vars1e-5.FULL_LIST.1e-5.beta.tsv") %>% drop_na() %>% filter(ids %in% z_scores$ids) %>% arrange(ids) %>% select(-ids) %>% as.matrix() 
all_ids <- fread("~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/seed2_thresh0.9_h2-0.1_vars1e-5.FULL_LIST.1e-5.beta.tsv") %>% drop_na() %>% filter(ids %in% z_scores$ids) %>% arrange(ids) %>% select(ids)
W_ones <- matrix(1,nrow = nrow(W_maf), ncol = ncol(W_maf))
```

## Set up parameters for factorization....
```{r}
option <- list()
 option[['K']] <- 15
option[['alpha1']] <- 1
option[['lambda1']] <- 1
option[['iter']] <- 20
option[['ones_mixed_std']] <- FALSE
option[['ones_mixed_ref']] <- FALSE
option[['ones']] <- FALSE
option[['disp']] <- FALSE
option[['convF']] <- 0
option[['convO']] <- 0
option[['ones_mixed_std']] <- FALSE
option[["ones_mixed_ref"]] <- FALSE
option[['ones_eigenvect']] <- TRUE
option[['ones_plain']] <- FALSE
option[['reweighted']] <- FALSE
option[["glmnet"]] <- FALSE
option[["parallel"]] <- TRUE
#Store stats here
run_stats_l1 <- list()
run_stats_flashr <- list()
```
### First test: Regular z
```{r}
option[['alpha1']] <- 10
option[['lambda1']] <- 10
option[["parallel"]] <- TRUE
W <- W_ones
X <- z
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$regular_z_parallel <- c(run, time)
```
The above parallelized ran in Time difference of 4.696587 mins. Viz this
```{r}
plotFactors(run_stats_l1$regular_z_parallel[[1]], trait_names = names, title = "Regular Z, no scaled, parallel")
```
Run in regular:
```{r}
option[["parallel"]] <- FALSE
option[["fastReg"]] <- FALSE
W <- W_ones
X <- z
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$regular_z <- c(run, time)
```
The above ran in 8.777028 mins. Now viz it:
```{r}
plotFactors(run_stats_l1$regular_z[[1]], trait_names = names, title = "Regular Z, no scaled, serial")
```
More sparse:
```{r}
option[['alpha1']] <- 17
option[['lambda1']] <- 17
option[["parallel"]] <- TRUE
option[["fixed_ubiq"]] <- FALSE
W <- W_ones
X <- z
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$regular_z_parallel_sparser <- c(run, time)
#load("~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/l1_with_ubiq_factor/weighting_runs.RData")


```


aaalright. I'm too impatient. Let's try glmnet.
```{r}
option[["glmnet"]] <- TRUE
option[['alpha1']] <- 0.1
option[['lambda1']] <- 0.1
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$regular_z_glmnet <- c(run, time)
plotFactors(run[[1]],trait_names = names, title = "Regular Z, glmnet")
```
And then with flash R
```{r}
option[["glmnet"]] <-FALSE
start <- Sys.time()
run <- flashr:::flash_greedy_workhorse(X,
                                          var_type = "constant",
                                          ebnm_fn = "ebnm_ash",
                                          Kmax = 15,
                                          verbose_output = "odF")
end <- Sys.time()
time <- end-start
run_stats_l1$regular_z_flashr <- c(run, time)
out <- flash_get_ldf(run)
plotFactors(out$f, trait_names = names, title = "Regular Z, flashR")
```
Def seems to be slow on startup. Wonder why? why all the run things?
Okay, so far glmnet seems to be waaay slower. Not as promising as I was hoping.
The solution I think is to parallelize the regression on the loading matrix. If we can speed that up, we are in buisness.

### Second test: Betas alone
I guess the scale of the betas requires smaller reg params
```{r}
option[["glmnet"]] <- FALSE
option[['alpha1']] <- 0.01
option[['lambda1']] <- 0.01
option[["parallel"]] <- TRUE
W <- W_ones
X <- betas
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$betas_parallel <- c(run, time)
plotFactors(run[[1]],trait_names = names, title = "Regular B, parallel")
```


### Third test: Betas, with SE
check things
```{r}
option[['alpha1']] <- 0.00001
option[['lambda1']] <- 0.00001
W <- W_se
X <- betas
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$betas_se_parallel <- c(run, time)
plotFactors(run_stats_l1$betas_se_parallel[[1]],trait_names = names, title = "B, W = SE, parallel")
```
Okay, this looks really weird to me. Heavies loading on spherical power right? What is going on here?
Let's quickly take a look at these traits, the number of samples they have, their heritabilities, etc. and see what is going on.
Based on what I see, it basically ranks them by their heritabilities! nice.
```{bash}
source ~/.bash_profile
downloadFromBluecrab /work-zfs/abattle4/ashton/snp_networks/gwas_decomp_ldsc/trait_data_seed2_thresh0.9_h20.1.txt ~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/
```

COME BACK AND CONFIRM THIS. BUT THIS IS WHAT I SEE....
First- try just the raw f1 loadings....
```{r}
load("~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/l1_with_ubiq_factor/weighting_runs.RData")
trait_data <- fread("~/Documents/JHU/Research/LocalData/snp_network/seed2_thresh0.9_h2-0.1_vars1e-5/trait_data_seed2_thresh0.9_h20.1.txt")
data_order <- colnames(z_scores)[-1]
trait_data <- trait_data %>% mutate("order" = factor(phenotype, levels = data_order)) %>% arrange(order)  %>% mutate("matrix_f1" = abs(run_stats_l1$betas_se_parallel[[1]][,1]))
trait_data %>% select(phenotype, description, h2_liability, h2_liability_se, matrix_f1, n) %>% arrange(-matrix_f1)
cor(trait_data$h2_liability, trait_data$matrix_f1)
cor(trait_data$h2_liability_se, trait_data$matrix_f1)
cor(trait_data$h2_observed, trait_data$matrix_f1)
cor(trait_data$h2_observed_se, trait_data$matrix_f1)
cor(trait_data$n, trait_data$matrix_f1)
colnames(trait_data)
```

```{r}
trait_data <- trait_data %>% mutate("order" = factor(phenotype, levels = data_order)) %>% arrange(order)  %>% mutate("matrix_f1" = abs(run_stats_l1$betas_se_parallel[[1]][,1]))
trait_data %>% select(phenotype, description, h2_liability, h2_liability_se, matrix_f1, n) %>% arrange(-matrix_f1)
cor(trait_data$h2_liability, trait_data$matrix_f1)
cor(trait_data$h2_liability_se, trait_data$matrix_f1)
cor(trait_data$h2_observed, trait_data$matrix_f1)
cor(trait_data$h2_observed_se, trait_data$matrix_f1)
cor(trait_data$n, trait_data$matrix_f1)
colnames(trait_data)
```
Let's visualize this in a useful way....
```{r}
toscale <- trait_data %>% select(phenotype, description, h2_liability, h2_liability_se, matrix_f1, n, h2_z, lambdaGC, confidence)
scaled <- data.frame(sapply(3:ncol(toscale), function(x) toscale[,..x] / max(toscale[,..x]))) %>% mutate("phenotype" = toscale$phenotype, "description" = toscale$description) %>% pivot_longer(cols = c(h2_liability, h2_liability_se, matrix_f1, n, h2_z, lambdaGC))
ggplot(data =scaled, aes(x = name, y = phenotype, fill = value)) + geom_tile()
```

```{r}
tocor <- as.matrix(trait_data %>% select(h2_liability, h2_liability_se, matrix_f1, n, h2_z, lambdaGC, intercept, mean_chi2))
Hmisc::rcorr(tocor)
trait_data
```

Okay. So it looks likewe are really picking up on the lambdaGC.
How about when we just do PCA- what is the top PC picking up on?
```{r}
pcs <- prcomp(z_scores[,-1],center = T, scale. = T)
evect <- pcs$rotation[,1]
evect_t <- pcs$rotation[1,]
pc <- pcs$x[1,]

Hmisc::rcorr(tocor, pc)
Hmisc::rcorr(tocor, evect)
Hmisc::rcorr(tocor, evect_t)
```
As a matter of comparison...... svd?
```{r}
pcs <- prcomp(z_scores[,-1],center = T, scale. = T)
svd <- svd(scale(z_scores[,-1]))
evect <- svd$v[,1]
Hmisc::rcorr(tocor, evect)
```

So specifically, corresponds with n and lambda GC
```{r}
summary(lm(data= data.frame(tocor), n ~ matrix_f1))
summary(lm(data= data.frame(tocor), lambdaGC ~ matrix_f1))
```
Okay, this looks really weird to me. Heavies loading on spherical power right? What is going on here?
Let's quickly take a look at these traits, the number of samples they have, their heritabilities, etc. and see what is going on.
Based on what I see, it basically ranks them by their heritabilities! nice.
Not quuite the case as the above revealed.
Less spasre cuz yea.
```{r}
option[['alpha1']] <- 1e-7
option[['lambda1']] <- 1e-7
option[["fixed_ubiq"]] <- FALSE
W <- W_se
X <- betas
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$betas_se_parallel_unsparse <- c(run, time)
plotFactors(run[[1]],trait_names = names, title = "B, W = SE, parallel, param = 1e-7")

save(run_stats_l1, file = "~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/l1_with_ubiq_factor/weighting_runs.RData")
```
This really doesn't seem to be working so well, does it....
```{r}
option[['alpha1']] <- 1e-7
option[['lambda1']] <- 1e-6
option[["fixed_ubiq"]] <- FALSE
W <- W_se
X <- betas
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$betas_se_parallel_unsparse <- c(run, time)
plotFactors(run[[1]],trait_names = names, title = "B, W = SE, parallel, param = 1e-7")

save(run_stats_l1, file = "~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/l1_with_ubiq_factor/weighting_runs.RData")
```
The way these plots are looking, I wonder if my SE data isn't wrong somehow...
Try not parallel?
```{r}
option[['alpha1']] <- 1e-7
option[['lambda1']] <- 1e-6
option[["parallel"]] <- FALSE
option[["fastReg"]] <- FALSE
W <- W_se
X <- betas
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$betas_se_notparallel_unsparse <- c(run, time)
plotFactors(run[[1]],trait_names = names, title = "B, W = SE, not parallel, param = mixed")

save(run_stats_l1, file = "~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/l1_with_ubiq_factor/weighting_runs.RData")
```



### Fourth test: Betas, with var MAF
This should basically favor more common variants over rare ones.
```{r}
option[['alpha1']] <- 1e-3
option[['lambda1']] <- 1e-3
option[["fixed_ubiq"]] <- FALSE
W <- W_varmaf
X <- betas
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$betas_maf_parallel <- c(run, time)
#plotFactors(run[[1]],trait_names = names, title = "B, W = var(MAF), parallel, param = 1e-6")

save(run_stats_l1, file = "~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/l1_with_ubiq_factor/weighting_runs.RData")
```
Some plots
```{r}
plotFactors(run_stats_l1$betas_maf[[1]], trait_names = names, title = "MAF corrected Beta")
names(run_stats_l1)
```

order the right way....
```{r}
dist <- hclust(as.dist(1-cor(X)))
orderh <- dist$order
plotFactors(run_stats_l1$betas_parallel[[1]], trait_names = names, title = "Betas only",cluster = F, t_order = orderh)
plotFactors(run_stats_l1$betas_maf_parallel[[1]], trait_names = names, title = "Betas, W = var(MAF)",cluster = F, t_order = orderh)
plotFactors(run_stats_l1$betas_se_parallel_unsparse[[1]], trait_names = names, title = "Betas, W = SE",cluster = F, t_order = orderh)
plotFactors(run_stats_l1$betas_se_notparallel_unsparse[[1]], trait_names = names, title = "Betas, W = SE (serial run)",cluster = F, t_order = orderh)
plotFactors(run_stats_l1$betas_se_LL2[[1]], trait_names = names, title = "Betas, W = SE, L2 on L", cluster = F, t_order = orderh)
plotFactors(run_stats_l1$regular_z_parallel[[1]], trait_names = names, title = "Z only, W = 1",cluster = F, t_order = orderh)
plotFactors(run_stats_l1$regular_z_parallel_sparser[[1]], trait_names = names, title = "Z only, W = 1, sparser",cluster = F, t_order = orderh)


```



IDEA; what if I ask for L2 regularization on the loadings matrix? Let that be more continuous, but require the factors to be sparse.
```{r}
option[['alpha1']] <- 1e-7
option[['lambda1']] <- 1e-6
option[["parallel"]] <- FALSE
option[["ridge_L"]] <- TRUE
W <- W_se
X <- betas
start <- Sys.time()
run <- Update_FL(X, W, option)
end <- Sys.time()
time <- end-start
run_stats_l1$betas_se_LL2 <- c(run, time)
plotFactors(run[[1]],trait_names = names, title = "L2 on loadings")

save(run_stats_l1, file = "~/Documents/JHU/Research/LocalData/snp_network/sn-spMF/ts_eQTLs/l1_with_ubiq_factor/weighting_runs.RData")
```